A Bayesian search to find \\high-mass black holes in LIGO data

### 2020-08-21 main.tex
improve introduction

If intermediate-mass black holes are present, the gravitational waves emitted from the merger of a binary black hole system with at least one intermediate-mass black hole (up to a mass of $400$ \msun) will be detectable by the ground-based gravitational-wave network. Gravitational waves from such systems should occur at a rate of $0-10\text{yr}^{-1}$ \cite{fregeau2006imbhbRatePrediction, mandel2008rates,rodriguez2015bbhRatePredictions}.  However, even after conducting a targeted match-filter based search for gravitational waves from intermediate-mass black holes the largest total mass detected so far is approximately $80$ \msun \cite{salemi2019search, abbott2019gwtc}. A possible explanation for the absence of intermediate-mass events may be due to their misclassification as short-duration instrumental noise transients \cite{blip_glitches}. These glitches can mimic astrophysical signals and hence decrease the significance of true gravitational wave events. \\

% Launch into what Bayesain means

% mathced-filter searches do not utilise as much information as present
% an optimal search would computer a bayesian odds
% rest of para what a bayesian odds is 
% typicallly Bayesain a follow up tool
% no reason why we cnat use for search
% bayesian search on an entire observing run

One method to account for glitches while searching for gravitational waves from coalescing compact binaries is by utilising a signal consistency test \cite{pycbc_short_duration_transients} or a signal coherence test amongst a network of detectors \cite{bci, bcr_paper,  bcr_gw151216, bayesian_odds}. A signal-coherence test such as the Bayesian coherence ratio (\bcr) \cite{bcr_paper} provides a Bayesian odds ratio comparing the probability that the data contains an astrophysical signal versus the probability that the data contains an incoherent glitch or noise. Although an odds ratio, the \bcr can be used  as a frequentistic \textit{ranking statistic}, ranking the likelihood that a trigger is due to a gravitational-wave signal \cite{bcr_paper}. The \bcr as a ranking statistic is much more powerful than a matched-filter SNR as the \bcr incorporates numerous features that may distinguish gravitational waves from noise and demands that signals must be coherent while glitches must be incoherent amongst different detectors. \\

 In this paper we utilise the \bcr to calculate \pastro, the probability that a signal is of astrophysical origin [cite],  for all the high-mass (systems with total masses in the range of $55-500$ \msun) candidate gravitational-wave signals from compact binary coalescences in the detector data recorded during aLIGO's second observing run (O2).  We find that (a) our search does not identify any new stellar mass or intermediate mass black holes; (b) high-mass events reported in the GWTC-1, including GW170729 (an event with disputed $p_\text{astro}$ amongst various search pipelines) have high \pastro; and that (c) high-mass events detected from independent groups have low \pastro. \\

The remainder of this paper is structured as follows. We outline our analysis methods in Section~\ref{sec:method}. We present our results in Section~\ref{sec:result}, and discuss these results in the context of the significance of gravitational wave candidates in Section~\ref{sec:Conclusion}.

@@

@@

To calculate the Bayesian evidences $Z^S$, $Z_{Gi}$ and $Z_{Ni}$, we carry out Bayesian inference with \bilby CITE, using \dynesty CITE as our nested sampler.

>>>>

If intermediate-mass black holes are present, the gravitational waves emitted from the merger of a binary black hole system with at least one intermediate-mass black hole (up to a mass of $400$ \msun) will be detectable by the ground-based gravitational-wave network. Gravitational waves from such systems should occur at a rate of $0-10\text{yr}^{-1}$ \cite{fregeau2006imbhbRatePrediction, mandel2008rates,rodriguez2015bbhRatePredictions}.  However, even after conducting a targeted matched-filter based search for gravitational waves from intermediate-mass black holes the largest total mass detected so far is approximately $80$ \msun \cite{salemi2019search, abbott2019gwtc}. A possible explanation for the absence of intermediate-mass events may be due to their misclassification as short-duration instrumental noise transients known as glitches \cite{glitch_in_fifth_ligo_run, bayeswave, improving_dq_in_early_runs, ligo_glitch_gw150914, pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. These glitches can mimic astrophysical signals and hence decrease the significance of true gravitational wave events. \\

One method to account for glitches while searching for gravitational waves from coalescing compact binaries is by utilising an astrophysical Bayesian odds  \cite{bci, BCR1, BCR2, bcr_gw151216, bayesian_odds}. A true Bayesian odds calculated without using bootstrap techniques can provide events with a more accurate significance that is agnostic to a specific search strategy \cite{BCR2, bcr_gw151216,  bayesian_odds}. Additionally, a Bayesian odds can include information such as if the gravitational wave event's signal is coherent amongst the network of detectors, if the binary system that created the gravitational waves was precessing, and if the gravitational wave signal contains higher-order modes.  It is because Bayesian methods can incorporate all this physical information about a gravitational wave signal that the LIGO Scientific Collaboration uses these methods to determine the source parameters of gravitational wave events \cite{abbott2016ligo, abbott2019gwtc}. This paper demonstrates that the power of Bayesian methods used in parameter estimation can also successfully be used to discriminate between coherent gravitational-wave signals, incoherent glitches, and Gaussian noise in the form of a Bayesian search. \\

 In this paper we utilise a Bayesian method to search for the significant gravitational wave signals from high-mass (systems with total masses in the range of $55-500$ \msun) compact binary coalescences in the detector data recorded during aLIGO's second observing run (O2).  We find that (a) our search does not identify any unreported stellar mass or intermediate mass black holes; (b) high-mass events reported in the GWTC-1, including GW170729 (an event with disputed $p_\text{astro}$ amongst various search pipelines) have high low significance; and that (c) high-mass events detected from some independent groups have low significance. \\

The remainder of this paper is structured as follows. We outline our analysis methods, including details of our Bayesian  in Section~\ref{sec:method}. We present our results in Section~\ref{sec:result}, and discuss these results in the context of the significance of gravitational wave candidates in Section~\ref{sec:Conclusion}.

@@

\pastro, the probability that a signal is of astrophysical origin \cite{pastro_1,pastro_2,pastro_3},

a signal consistency test \cite{pycbc_short_duration_transients} or a signal coherence test amongst a network of detectors. A signal-coherence test such as the Bayesian coherence ratio (\bcr) \cite{bcr_paper} provides a Bayesian odds ratio comparing the probability that the data contains an astrophysical signal versus the probability that the data contains an incoherent glitch or noise. Although an odds ratio, the \bcr can be used  as a frequentistic \textit{ranking statistic}, ranking the likelihood that a trigger is due to a gravitational-wave signal \cite{bcr_paper}. The \bcr as a ranking statistic is much more powerful than a matched-filter SNR as the \bcr incorporates numerous features that may distinguish gravitational waves from noise and demands that signals must be coherent while glitches must be incoherent amongst different detectors. \\

@@

To calculate the Bayesian evidences $Z^S$, $Z_{Gi}$ and $Z_{Ni}$, we carry out Bayesian inference with \bilby \cite{bilby}, using \dynesty \cite{dynesty} as our nested sampler.
